Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-13 paper.tex
anthropomorphism definitions etc.

\vspace{2cm}
\section{Introduction and Overview}
\label{sec:intro}

%DEFINITION OF ANTHROPOMORPHISM

The essence of anthropomorphism is the perception of human-like characteristics
in either real or imagined non-human agents. According to Epley \textit{et al.}
\cite{epley_when_2008}, these human-like characteristics may include physical
appearance, emotional states perceived to be uniquely human, or inner mental
states and motivations. Real or imagined non-human agents can be "anything that
acts -- or is believed to act -- with apparent independence, including non-human
animals, natural forces, religious agents, technological gadgets, or mechanical
devices". Such anthropomorphic representations are important determinants of how
a person perceives and in turn behaves towards these agents, or how a person may
behave in light of these agents \cite{epley_when_2008}.

%ANTHROPOMORPHISM AS SUBJECT IN HCI AND HRI

Phenomenons related to anthropomorphism -- along with questions such as "how
much human do we perceive in the non-human" or "how human-like do we treat
non-human artifacts" (and why?) -- have received remarkable attention in the
domains of Human-Computer Interaction (HCI) and Human-Robot Interaction (HRI):
the study of humans interacting with an artificial system / machine. Following
the \textit{Media Equation}, it has been described that people mindlessly react
socially to computers \cite{reeves_media_1996} and may even attribute human
(social) characteristics to their artificial counterpart. The phenomenon has
also been observed with virtual characters \hl{cite [50, 100] in rosenthal-von
der P{\"u}tten (2013)}, and more recent studies investigated how far people
ascribe personality or intentions to artificial embodied agents or whether they
show empathetic behavior toward robots
\cite{rosenthal-vonderputten_experimental_2013}. The common base for this
research is the assumption that humans perceive or react to a non-human agent or
artifact as somewhat similar to a human.

%CONTROVERSAL UNDERSTANDINGS OF ANTHROPOMORPHISM

However, not everything colloquially labeled as \emph{anthropomorphism} really
is anthropomorphism. Different, even partially conflicting, understandings of
anthropomorphism exist across and within disciplines
\cite{duffy_anthropomorphism_2002}. Epley \textit{et al.} \cite{epley_when_2008}
try to narrow down and clarify the understanding of anthropomorphism in
psychology by explaining what it is \textit{not}. The authors list four things
that anthropomorphism does not include: First, anthropomorphism does not include
behavioral descriptions of observable actions. Anthropomorphism requires going
beyond what is directly observable. Second, anthropomorphism does not merely
entail animism, as animate life is not a uniquely human-like characteristic.
Third, anthropomorphism does not include any requirement of reasoned or
reflective endorsement of an inference (\hl{the strength of anthropomorphic
inferences can vary} \footnote{for further explanation the reader may refer to
the original work \cite{epley_when_2008}.}).	Finally, anthropomorphism is not
necessarily inaccurate. A psychological theory should take this (in)accuracy in
anthropomorphism into account and predict variability in the tendency to
perceive human-like traits in non-human agents.  In conclusion, it has been
noted that anthropomorphism is a very complex phenomenon - in itself, its
understanding, and when studying it. A common formal ground is missing; however,
it is important to be clear about what exactly we mean when using the term
anthropomorphism.

@@

\vspace{2cm}
\section{Anthropomorphism}
\label{sec:anthropomorphism}

@@

The term \textbf{anthropomorphism} is used in different senses throughout
natural sciences, psychology, Human-Computer Interaction, and Human-Robot
Interaction \cite{duffy_anthropomorphism_2003}. Originally, anthropomorphism
comes from the Greek \textit{`anthropos'} for `man' (or `human') and
\textit{`morphe'} for `form/structure' (or `shape'). It can be understood as
people's tendency to think of objects as if they had human characteristics, and
consequently, attribute / ascribe specific human characteristics to these
non-human entities and artifacts (including animals, and robots)
\cite{duffy_anthropomorphism_2003,schmitz_concepts_2011}. Apart from
anthropomorphism there is also \textbf{animism}, the \textit{"attribution of
life to the nonliving"} \hl{in Schmitz 2011, cite Piaget original text!}. Often
synonymously used to animism is \textbf{zoomorphism}, which describes the case
when non-lifelike objects are associated with animalistic attributes, excluding
human-specific traits. Both anthropomorphism and zoomorphism can be embraced by
the concept of \textbf{life-likeness}. \footnote{The inverse process would be
called dehumanization \cite{haslam_dehumanization:_2006}, or mechanomorphism
\cite{caporael_anthropomorphism_1986}.}

@@

>>>>

\section{Introduction and Overview}
\label{sec:intro}

The essence of anthropomorphism is the perception of human-like characteristics in either real or imagined non-human agents. In Human-Robot Interaction (HRI), anthropomorphism is a commonly observed phenomenon: people tend to ascribe human-like characteristics to robots, and they are willing to treat and react to robots as if they were human-social counterparts. The HRI and social robotics literature is full of various kinds of studies showing how humans readily interact in a social fashion with robots -- despite the fact that robots are not humans. In our own studies on HRI analysis in real-world settings, we also observed anthropomorphism. However, we found qualitative differences and dynamics over time in how participants anthropomorphized and related to the robots used in our studies. More concretely, we found that anthropomorphism was not only based on how much the robot resembles a human (anthropomorphic design) but that it was determined by multiple factors, including individual characteristics of the human user, and the context of use, for instance. We also found that anthropomorphism changed over time with growing experience of the user with the system. By trying to understand these variances and dynamics in anthropomorphism, we reviewed related work from the HRI domain and social robotics. Despite the fact that there is a vast amount of literature on anthropomorphism in HRI, we were surprised to find that other studies do not describe similar variances or dynamics. Instead, we came across diverse and partially conflicting understandings of anthropomorphism. We also had the impression that the reported data were sometimes over-interpreted (probably due to the confusion of what exactly anthropomorphism is) or obtained in settings and assessed in ways that we believe do not give enough consideration to the complex nature of anthropomorphism.\footnote{As anthropomorphism emerges from an interaction between a human and a non-human agent, we believe that showing pictures of different kinds of robots and then using questionnaires to measure anthropomorphism is somehow limited. In the discussion in section \ref{sec:discussion} we discuss what tools and measurements could be used to study anthropomorphism and its dynamics in interaction studies in real-world environments.} However, when reviewing literature from social psychology, cognitive science, and neuroscience, we found support for the variances and dynamics in anthropomorphism that we observed in our studies. As HRI integrates many disciplines, we think that it is valid to use theories from these related domains for interpreting our results and that it is relevant to report on this integration.

In this article, we would like to share some critical but constructive considerations on anthropomorphism in robotics. Thereby, we contribute to the understanding of this complex phenomenon related to HRI. In accordance with theories from other related domains, we propose that anthropomorphism is determined and affected by multiple factors, and we show how this can impact the user experience and acceptance of robots in daily life environments. We further suggest an initial model that we call ``dynamics of anthropomorphism'' that serves to describe how anthropomorphism evolves and shapes the interaction over time. We illustrate our idea with empirical findings from several HRI studies that we conducted in natural environments. We then present a cognitive and neuroscience perspective on anthropomorphism to consolidate our considerations and the observed dynamics. Finally, we discuss alternative ways to measure anthropomorphism in HRI, as this is a critical aspect of gaining a better understanding of it.

@@

\section{Anthropomorphism}
\label{sec:anthropomorphism}

In this section we will clarify the terminology that we use as we have found conflicting understandings of the term ``anthropomorphism'' in the related literature. We first present anthropomorphism as a social phenomenon. Then we propose to use another term, namely ``anthropomorphic form'', to refer to the human-inspired design of artifacts. Finally, we focus on anthropomorphism in Human-Robot Interaction, as it naturally brings together (and needs to question) the usage of an anthropomorphic design of robots and its consequences for the interaction with the human user.

@@

Anthropomorphism is an interesting phenomenon because it tells us something about the relationship between the human who anthropomorphizes, and the non-human agent which is anthropomorphized. Though commonly observed, it appears controversial, at the same time, that humans tend to perceive human-like physical appearance (\eg faces), emotional states, or inner mental states and motivations in non-human agents such as animals, natural forces, religious agents, technological gadgets, or mechanical devices \citep{epley_when_2008}.\footnote{According to \cite{epley_when_2008}, a non-human agent can be anything that acts -- or is believed to act -- with apparent independence.}  Anthropomorphism has received remarkable attention across several domains, such as natural sciences, psychology, Human-Computer Interaction, and Human-Robot Interaction. However, the term is used in different senses throughout these disciplines \citep{duffy_anthropomorphism_2003} which makes it difficult to synthesize findings.

Originally, \textbf{anthropomorphism} comes from the Greek \textit{`anthropos'} for `man' (or `human') and \textit{`morphe'} for `form / structure' (or `shape'). It can generally be understood as people's tendency to perceive a non-human agent as if it had some human characteristics. In turn, how a person perceives an agent further determines how she relates to and behaves towards the agent, or how she may behave in light of this agent \citep{epley_when_2008}. By this, anthropomorphism can be seen as a special kind of relationship between the human and the non-human agent. This relationship is often characterized by the fact that the human attributes or ascribes uniquely human characteristics to the non-human entity or artifact. The most prominently anthropomorphized non-human agents are probably animals and robots \citep{duffy_anthropomorphism_2003,schmitz_concepts_2011}.

Apart from anthropomorphism there exists also \textbf{animism}, which can be understood as the \textit{"attribution of life to the nonliving"} \hl{in Schmitz 2011, cite Piaget original text!}. Using this definition, animism differs in two main aspects from anthropomorphism:
First, the property which is ascribed differs: it is not human-unique but applies to all living beings (life-likeness). For instance, attributing life-likeness to a stone must not necessarily be labeled as anthropomorphism, since being alive is not a human unique property but applies to all living beings.
The second difference comes from the instance (or better things) to which life is attributed: all living entities are excluded in this definition. In other words, pets as non-human but living beings can only be anthropomorphized (because they are not human) but not animized (because they are already alive).
Often synonymously used to animism is the term \textbf{zoomorphism}, which applies to the cases when non-lifelike instances / things are associated with animalistic attributes, excluding human-specific traits. Both anthropomorphism and zoomorphism can be embraced by the concept of \textbf{life-likeness}. \footnote{The inverse process would be called dehumanization \cite{haslam_dehumanization:_2006}, or mechanomorphism \cite{caporael_anthropomorphism_1986}.}

@@

Following research in Human-Computer Interaction (HCI) the \textit{Media Equation} \citep{reeves_media_1996} describes that people mindlessly react socially to their computer and may even attribute human (social) characteristics to it. The phenomenon has also been observed with virtual characters, and artificial embodied agents, such as robots. For instance, recent studies found that people tend to ascribe personality and intentions to robots \hl(REFS) and that they show empathetic behavior toward robots \cite{rosenthal-vonderputten_experimental_2013}. To subsume it, there is a growing body of literature that describes anthropomorphism in HCI and HRI.

However, not everything colloquially labeled as \emph{anthropomorphism} really is anthropomorphism. Different, even partially conflicting understandings of anthropomorphism exist across and within disciplines \cite{duffy_anthropomorphism_2002}. \cite{epley_when_2008} try to narrow down and clarify the understanding of anthropomorphism in psychology by explaining what it is \textit{not}. The authors list four things
that anthropomorphism does not include: First, anthropomorphism does not include
behavioral descriptions of observable actions. Anthropomorphism requires going
beyond what is directly observable. Second, anthropomorphism does not merely
entail animism, as animate life is not a uniquely human-like characteristic.
Third, anthropomorphism does not include any requirement of reasoned or
reflective endorsement of an inference (\hl{the strength of anthropomorphic
inferences can vary} \footnote{for further explanation the reader may refer to
the original work \cite{epley_when_2008}.}).	Finally, anthropomorphism is not
necessarily inaccurate. A psychological theory should take this (in)accuracy in
anthropomorphism into account and predict variability in the tendency to
perceive human-like traits in non-human agents.  In conclusion, it has been
noted that anthropomorphism is a very complex phenomenon - in itself, its
understanding, and when studying it. A common formal ground is missing; however,
it is important to be clear about what exactly we mean when using the term
anthropomorphism.
